### 2019-05-03 tex_v4/manuscript.tex
post-defense, corrections and revisions

\begin{document}
	\begin{abstract}
		The acorn weevil (\textit{Curculio} Linnaeus, 1758) rostrum (snout) exhibits remarkable flexibility and toughness derived from the microarchitecture of its exoskeleton.
		Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and fracture toughness of the distal portion of the snout.
		Using Classical Laminate Plate Theory, we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
		We show that the tensile behavior of the rostrum across six \textit{Curculio} species with high morphological variation correlates with changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton.
		Accordingly, increased endocuticle thickness is strongly correlated with increased tensile strength.
		Rostrum stiffness is shown to be inversely correlated with fracture toughness; thus allowing a highly curved rostrum to completely straighten without structural damage.
		Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure, and as a source of morphological constraint on the evolution of the rostrum in \textit{Curculio} weevils.

	\end{abstract}

@@

		The specimen did not show any evidence of fractures, micro-tears, or shear cusps anywhere in the surface of the exocuticle.
		Moreover, the tensile strength of the tested specimen's rostrum was consistent with that of other species members ($F_max~=~5.02~\,\text{N}$).
		Surprisingly, the specimen remained undamaged by the testing.
		We therefore conclude that under normal life conditions, repeated bending of the rostrum does not exceed the yield strength of the cuticle.

>>>>

\begin{document}
	\begin{abstract}
		The acorn weevil (\textit{Curculio} Linnaeus, 1758) rostrum (snout) exhibits remarkable flexibility and toughness derived from the microarchitecture of its exoskeleton.
		Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and toughness of the distal portion of the snout.
		Using Classical Laminate Plate Theory, we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
		We show that the tensile behavior of the rostrum across six \textit{Curculio} species with high morphological variation correlates with changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton.
		Accordingly, increased endocuticle thickness is strongly correlated with increased tensile strength.
		Rostrum stiffness is shown to be inversely correlated with work of fracture; thus allowing a highly curved rostrum to completely straighten without structural damage.
		Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure, and as a source of morphological constraint on the evolution of the rostrum in \textit{Curculio} weevils.
	\end{abstract}

@@

		The specimen did not show any evidence of fractures, micro-tears, or shear cusps anywhere in the surface of the exocuticle.
		Moreover, the tensile strength of the tested specimen's rostrum was consistent with that of other species members ($F_{max}~=~5.02~\,\text{N}$).
		Surprisingly, the specimen remained undamaged by the testing.
		We therefore conclude that under normal life conditions, repeated bending of the rostrum does not exceed the yield strength of the cuticle.

### 2019-05-03 tex_v5/manuscript.tex
post-defense, corrections and revisions

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{textgreek}
\usepackage{xcolor}
\usepackage[colorlinks]{hyperref}
	\hypersetup{colorlinks,
	linkcolor={red!50!black},
	citecolor={blue!60!black},
	urlcolor={blue!40!black}
	}
\graphicspath{{figures/}}
%-------------------------------------------------------------------------------

@@

		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and fracture toughness (see Fig.~\ref{fig::tensile}).
		This implies that stiffer specimens -- and by extension, stiffer cuticle profiles -- are generally more brittle.
		We also observed a moderate, but not statistically significant, stiffening size-effect with respect to rostral length.
		This runs contrary to our expectation that a longer, more strongly curved rostrum would require increased flexibility to avoid fracture during oviposition \cite{Hughes2004eco, Bonal2011}.
		It is possible that longer rostra also have a longer transition gradient from basal to apical profile; thereby reinforcing the junction between the rostrum and head capsule against buckling.
		Young's modulus of the rostrum would be comparatively higher in such species, due to the higher volume fraction of exocuticle.
		We therefore posit that the gross elastic performance of the cuticle is consistent across the weevil genus \textit{Curculio}.
		A \emph{single} mechanism -- i.e., the modified composite profile -- likely confers increased flexibility and tensile strength to the rostral apex across all \textit{Curculio} species.
		In addition, the endocuticle demonstrably contributes more to rostral tensile strength than the exocuticle; likely because of its organization into large bundles of aligned, anisotropic fibers, and amounting a trade-off between rigidity and toughness.
		Consequently, the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture-resistant.

@@

		Our results imply that fracture initiation occurs in the comparatively brittle exocuticle.
		The reduction in exocuticle thickness in the rostral apex might serve to mitigate crack formation in rostral bending.
		Based on this pattern of fracture behavior, we identified the exocuticle-rich occipital suture as a common point of void nucleation and crack initiation.
		From an evoluationary perspective, these findings reveal an unexpected morphological source of constraint on rostral flexibility, raising the intriguing possibility that this system evolved primarily via negative selection of fracture, rather than positive selection of flexibility.
		In particular, the cuticle is invaginated in precisely the portion of the snout that experiences the greatest degree of tension during antero-dorsal flexion.
		The doubly-thick exocuticle in the invagination thus creates an unavoidable, brittle weak-point in an otherwise endocuticle-dominated rostral apex.
		This constraint -- in conjunction with the minimization of exocuticle thickness in the rostral apex and the increased toughness derived from a thickened endocuticle -- lead us to consider that avoidance of catastrophic structural failure has been a driving selective pressure in the evolution of the female \textit{Curculio} rostrum.

@@

				\begin{equation}
					\begin{Bmatrix}
						\{N\} \\
						\{M\}
					\end{Bmatrix}
					=
					\begin{bmatrix}
						[A] & [B] \\
						[B] & [D]
					\end{bmatrix}
					\begin{Bmatrix}
					\{\epsilon^{\degree}\} \\
					\{\kappa\}
					\end{Bmatrix}\,.
				\end{equation}

@@

				\begin{equation}
				\label{eqn::abbd}
					\begin{aligned}
						\{N\} & = [A]\{\epsilon^{\degree}\}\,, \\
						\{M\} & = [D]\{\kappa\}\,,						
					\end{aligned}\,,
				\end{equation}

@@

				\begin{equation}
					\begin{aligned}
						\begin{Bmatrix}
							N_{xx} \\
							N_{yy} \\
							N_{xy}
						\end{Bmatrix}
						& =
						\begin{bmatrix}
							A_{11} & A_{12} & A_{16} \\
							A_{21} & A_{22} & A_{26} \\
							A_{61} & A_{62} & A_{66}
						\end{bmatrix}
						\begin{Bmatrix}
							\epsilon^{\degree}_{xx} \\
							\epsilon^{\degree}_{yy} \\
							\gamma^{\degree}_{xy}
						\end{Bmatrix}\,,
						\\
						\begin{Bmatrix}
							M_{xx} \\
							M_{yy} \\
							M_{xy}
						\end{Bmatrix}
						& =
						\begin{bmatrix}
							D_{11} & D_{12} & D_{16} \\
							D_{21} & D_{22} & D_{26} \\
							D_{61} & D_{62} & D_{66}
						\end{bmatrix}
						\begin{Bmatrix}
							\kappa_{xx} \\
							\kappa_{yy} \\
							\kappa_{xy}
						\end{Bmatrix}\,.
					\end{aligned}
				\end{equation}

@@

		\subsection*{Statistical analysis}
			\paragraph*{Model selection and fitting.}
				In order to explore the relationships between the composite structure and mechanical properties of the cuticle, we fit phylogenetic linear mixed-effects models (PGLMM) to the tensile testing data using maximum likelihood estimation \cite{Galecki2013, Hadfield2010, Houseworth2004, Stone2011glmm}.
				Raw data was processed using a custom script in Python version 3.5.2 \cite{Python}.
				Model exploration and fitting was conducted in \texttt{R version 3.5.1 (2018-07-02) -- ``Feather Spray''}, using the \texttt{`nlme'} and \texttt{`ape'} packages \cite{Rcite, Ape, Nlme}.
				Response variables and covariates were natural-log transformed, as needed, to ensure that the normalized model residuals were normally distributed [\texttt{R:shapiro.test}] and homoscedastic [\texttt{R:levene.test}], using numerical and graphical analysis (see Supplementary Data).
				In order to control for phylogenetic non-independence in the data, we included the species of each specimen as a random effect in all models.
				We also allowed for correlation in the error term of the models, as specified by a variance-covariance matrix generated from a Brownian motion model of trait evolution [\texttt{R:`ape':corBrownian}] along a phylogeny \cite{Munkemuller2012, Ape}.

				The preferred phylogeny for select North American \textit{Curculio} species is that of Bonal~et~al.~\cite{Bonal2016}.
				This phylogeny was generated using Maximum-Likelihood methods; it is untrametric yet has uniform internal branch lengths.
				The preferred tree was pruned to include only those species used in tensile testing, with a polytomy at each species' root node to represent the individual specimens examined for that species.
				Because the branch lengths of the tree were not specified, all branch lengths were set equal to 1.

				In general, phylogenetic regression is very robust with respect to missing or incomplete branch length data.
				Underestimation of branch lengths causes overestimation of phylogenetic signal.
				This phenomenon was not found to be statistically significant for any of our models.
				Therefore we consider our study's principal findings to be unaffected by branch length underestimation \cite{Munkemuller2012, Molina2017, Stone2011star}.

				Other models of trait evolution were considered during model exploration; including Ornstein-Uhlenbeck [\texttt{R:`ape':corMartins}] and variable rate (AC/DC) [\texttt{R:`ape':corBlomberg}] models \cite{Munkemuller2012, Ape}.
				However, neither of these produced a significantly better fit to the data than the Brownian motion model for any comparison, as measured by likelihood score and residual variance; viz. $R^2_{\sigma}$, and $R^2_{\beta^*}$ [\texttt{R:`r2glmm':r2beta}] (see Supplementary Data).
				$R^{2}_{\sigma}$ is the proportion of generalized variance explained by fixed effects.
				This measure is generally used for comparison of covariance structures \cite{Nakagawa2013, Jaeger2016, R2glmm}.
				$R^2_{\beta^*}$ measures the multivariate association between the outcome and the fixed effects within a given correlation structure, and is generally used to compare fixed effects \cite{Nakagawa2013, Jaeger2016, R2glmm}.
				Using the \texttt{R} package \texttt{`phytools'} \cite{Munkemuller2012, Phytools}, we also estimated the phylogenetic signal in the residual variance of each model; taking into account two measures of spatial autocorrelation (Abouheif's $\bar{C}$ and Moran's $I$), and two measures of phylogenetic signal (Pagel's $\lambda$ and Blomberg's $\kappa$).

				In all models, we tested whether the inclusion of phylogenetic correlation in the model error produced significantly better model fit, using a likelihood-ratio test [\texttt{R:`lmtest':lrtest}] and $R^{2}_{\sigma}$-difference test [\texttt{R:`r2glmm':r2beta(method=`sgv')}] between the fully-specified model and a model lacking the phylogenetic effect \cite{Lmtest, R2glmm}.
				Models that incorporated covariance due to Brownian motion consistently produced higher likelihood scores and fit to the data.
				However, no model exhibited statistically significant phylogenetic signal in any of the variables.
				Once models were fitted to the data, T- and F-statistics were calculated to determine whether each cofactor was significantly different from zero.

@@

				For our second aim we examined the relationship between specimen stiffness and resistance to fracture.
				Specimen stiffness was characterized using a low strain elastic modulus -- averaged across the first 33\% of the stress-strain curve -- and the secant modulus at failure.
				Resistance to fracture was quantified in terms of ultimate strain and fracture toughness, measured as area under the stress-strain curve.
				Four models were fitted, two of which had ultimate strain as the response variable, and two with fracture toughness as response variable.
				Each model then used one of the different measures of specimen stiffness as a fixed effect.

@@

			\bibitem{Houseworth2004}
				Housworth, E. A., Martins, E. P., and Lynch, M.
				The Phylogenetic Mixed Model
				\textit{Am. Nat.}
				\textbf{163:1,}
				84--96,
				(2004).

>>>>

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{textgreek}
\usepackage{xcolor}
\usepackage{microtype}
\usepackage[colorlinks]{hyperref}
	\hypersetup{colorlinks,
	linkcolor={red!50!black},
	citecolor={blue!60!black},
	urlcolor={blue!40!black}
	}
\graphicspath{{figures/}}
%-------------------------------------------------------------------------------

@@

		Instead, we found that the uniaxial elastic modulus (low strain: $E_{low}$) and secant modulus at failure ($E_{sec}$) were inversely correlated with ultimate strain and work of fracture (see Fig.~\ref{fig::tensile}).
		This implies that stiffer specimens -- and by extension, stiffer cuticle profiles -- are generally more brittle.
		We also observed a moderate, but not statistically significant, stiffening size-effect with respect to rostral length.
		This runs contrary to our expectation that a longer, more strongly curved rostrum would require increased flexibility to avoid fracture during oviposition \cite{Hughes2004eco, Bonal2011}.
		It is possible that longer rostra also have a longer transition gradient from basal to apical profile; thereby reinforcing the junction between the rostrum and head capsule against buckling.
		Young's modulus of the rostrum would be comparatively higher in such species, due to the higher volume fraction of exocuticle.
		We therefore posit that the gross elastic performance of the cuticle is consistent across the weevil genus \textit{Curculio}.
		A \emph{single} mechanism -- i.e., the modified composite profile -- likely confers increased flexibility and tensile strength to the rostral apex across all \textit{Curculio} species.
		In addition, the endocuticle demonstrably contributes more to rostral tensile strength than the exocuticle; likely because of its organization into large bundles of aligned, anisotropic fibers, and amounting a trade-off between rigidity and toughness.
		Consequently, the altered composite profile of the cuticle in the rostral apex makes the rostrum simultaneously more flexible and fracture-resistant.

@@

		Our results imply that fracture initiation occurs in the comparatively brittle exocuticle.
		The reduction in exocuticle thickness in the rostral apex might serve to mitigate crack formation in rostral bending.
		Based on this pattern of fracture behavior, we identified the exocuticle-rich occipital suture as a common point of void nucleation and crack initiation.
		From an evolutionary perspective, these findings reveal an unexpected morphological source of constraint on rostral flexibility, raising the intriguing possibility that this system evolved primarily via negative selection of fracture, rather than positive selection of flexibility.
		In particular, the cuticle is invaginated in precisely the portion of the snout that experiences the greatest degree of tension during antero-dorsal flexion.
		The doubly-thick exocuticle in the invagination thus creates an unavoidable, brittle weak-point in an otherwise endocuticle-dominated rostral apex.
		This constraint -- in conjunction with the minimization of exocuticle thickness in the rostral apex and the increased toughness derived from a thickened endocuticle -- lead us to consider that avoidance of catastrophic structural failure has been a driving selective pressure in the evolution of the female \textit{Curculio} rostrum.

@@

				\begin{equation}
					\begin{Bmatrix}
						\{N\} \\
						\{M\}
					\end{Bmatrix}
					=
					\begin{bmatrix}
						[A] & [B] \\
						[B] & [D]
					\end{bmatrix}
					\begin{Bmatrix}
					\{\epsilon^{\circ}\} \\
					\{\kappa\}
					\end{Bmatrix}\,.
				\end{equation}

@@

				\begin{equation}
				\label{eqn::abbd}
					\begin{aligned}
						\{N\} & = [A]\{\epsilon^{\circ}\}\,, \\
						\{M\} & = [D]\{\kappa\}\,,						
					\end{aligned}\,,
				\end{equation}

@@

				\begin{equation}
					\begin{aligned}
						\begin{Bmatrix}
							N_{xx} \\
							N_{yy} \\
							N_{xy}
						\end{Bmatrix}
						& =
						\begin{bmatrix}
							A_{11} & A_{12} & A_{16} \\
							A_{21} & A_{22} & A_{26} \\
							A_{61} & A_{62} & A_{66}
						\end{bmatrix}
						\begin{Bmatrix}
							\epsilon^{\circ}_{xx} \\
							\epsilon^{\circ}_{yy} \\
							\gamma^{\circ}_{xy}
						\end{Bmatrix}\,,
						\\
						\begin{Bmatrix}
							M_{xx} \\
							M_{yy} \\
							M_{xy}
						\end{Bmatrix}
						& =
						\begin{bmatrix}
							D_{11} & D_{12} & D_{16} \\
							D_{21} & D_{22} & D_{26} \\
							D_{61} & D_{62} & D_{66}
						\end{bmatrix}
						\begin{Bmatrix}
							\kappa_{xx} \\
							\kappa_{yy} \\
							\kappa_{xy}
						\end{Bmatrix}\,.
					\end{aligned}
				\end{equation}

@@

		\subsection*{Statistical analysis}
			\paragraph*{Model selection and fitting.}
				In order to explore the relationships between the composite structure and mechanical properties of the cuticle, we fit phylogenetic linear mixed-effects models (PGLMM) to the tensile testing data using maximum likelihood estimation \cite{Galecki2013, Hadfield2010, Housworth2004, Stone2011glmm}.
				Raw data was processed using a custom script in Python version 3.5.2 \cite{Python}.
				Model exploration and fitting was conducted in \texttt{R version 3.5.1 (2018-07-02) -- ``Feather Spray''}, using the \texttt{`nlme'} and \texttt{`ape'} packages \cite{Rcite, Nlme, Ape}.
				Response variables and covariates were natural-log transformed, as needed, to ensure that the normalized model residuals were normally distributed [\texttt{R:shapiro.test}] and homoscedastic [\texttt{R:levene.test}], using numerical and graphical analysis.
				In order to control for phylogenetic non-independence in the data, we included the species of each specimen as a random effect in all models.
				We also allowed for correlation in the error term of the models, as specified by a variance-covariance matrix generated from a Brownian motion model of trait evolution [\texttt{R:`ape':corBrownian}] along a phylogeny \cite{Munkemuller2012, Ape}.

				The preferred phylogeny for select North American \textit{Curculio} species is that of Bonal~et~al.~\cite{Bonal2016}.
				This phylogeny was generated using Maximum-Likelihood methods; it is untrametric yet has uniform internal branch lengths.
				The preferred tree was pruned to include only those species used in tensile testing, with a polytomy at each species' root node to represent the individual specimens examined for that species.
				Because the branch lengths of the tree were not specified, all branch lengths were set equal to 1.
				Underestimation of branch lengths causes overestimation of phylogenetic signal \cite{Munkemuller2012, Molina2017, Stone2011star}.
				We were unable to detect statistically significant phylogenetic signal for any of our models.
				For this reason we consider our study's principal findings to be unaffected by branch length underestimation.
				However, because of the relatively poor taxonomic resolution afforded by a sample of five species, we are unable to conclude whether significant phylogenetic signal exists in the traits examined herein.

				Other models of trait evolution were considered during model exploration; including Ornstein-Uhlenbeck [\texttt{R:`ape':corMartins}] and variable rate (AC/DC) models [\texttt{R:`ape':corBlomberg}] \cite{Munkemuller2012, Ape}.
				These models were chosen to compare among competing hypotheses of trait evolution, namely stabilizing selection and variable-rate evolution, respectively.
				We additionally tested a Brownian motion model of trait evolution, corresponding to variation constrained by clade membership (i.e., sister taxa are more similar to each other than to an out-group).
				However, neither of the former produced a significantly better fit to the data than the Brownian motion model for any comparison, as measured by likelihood score and residual variance; viz. $R^2_{\sigma}$, and $R^2_{\beta^*}$ [\texttt{R:`r2glmm':r2beta}].
				$R^{2}_{\sigma}$ is the proportion of generalized variance explained by fixed effects.
				This measure is generally used for comparison of covariance structures \cite{Nakagawa2013, Jaeger2016, R2glmm}.
				$R^2_{\beta^*}$ measures the multivariate association between the outcome and the fixed effects within a given correlation structure, and is generally used to compare fixed effects \cite{Nakagawa2013, Jaeger2016, R2glmm}.
				Using the \texttt{R} package \texttt{`phytools'} \cite{Munkemuller2012, Phytools}, we also estimated the phylogenetic signal in the residual variance of each model; taking into account two measures of spatial autocorrelation (Abouheif's $\bar{C}$ and Moran's $I$), and two measures of phylogenetic signal (Pagel's $\lambda$ and Blomberg's $\kappa$).

				In all models, we tested whether the inclusion of phylogenetic correlation in the model error produced significantly better model fit, using a likelihood-ratio test [\texttt{R:`lmtest':lrtest}] and $R^{2}_{\sigma}$-difference test [\texttt{R:`r2glmm':r2beta(method=`sgv')}] between the fully-specified model and a model lacking the phylogenetic effect \cite{Lmtest, R2glmm}.
				Models that incorporated covariance due to Brownian motion consistently produced higher likelihood scores and fit to the data.
				However, no model exhibited statistically significant phylogenetic signal in any of the variables.
				Our findings imply that \emph{residual variance} in the examined traits does not exhibit significant rate-variation or stabilizing selection in the amount or direction of residual variance between groups; rather, variation within and between clades is distributed randomly by clade, with limited phylogenetic dependence.
				Once models were fitted to the data, T- and F-statistics were calculated to determine whether each cofactor was significantly different from zero.

@@

				For our second aim we examined the relationship between specimen stiffness and resistance to fracture.
				Specimen stiffness was characterized using a low strain elastic modulus -- averaged across the first 33\% of the stress-strain curve -- and the secant modulus at failure.
				Resistance to fracture was quantified in terms of ultimate strain and work of fracture, measured as area under the stress-strain curve.
				Four models were fitted, two of which had ultimate strain as the response variable, and two with work of fracture as response variable.
				Each model then used one of the different measures of specimen stiffness as a fixed effect.

@@

			\bibitem{Housworth2004}
				Housworth, E. A., Martins, E. P., and Lynch, M.
				The Phylogenetic Mixed Model
				\textit{Am. Nat.}
				\textbf{163:1,}
				84--96,
				(2004).
